### 2014-01-31 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/abstract.tex
Draft abstract

A variety of conventional feature selection methods choose informative
features or feature sets and empirically lead to good accuracy
improvements, but few focus on feature selection w.r.t.\ specific loss
functions; in particular, there seems to be little work on feature
selection for classifiers if a secondary objective after accuracy is
either precision or recall.  In this paper we propose a novel
probabilistic model of feature selection based on a probabilistic
voting model and argue that MAP inference w.r.t. to two extreme
objectives in this model yield respective high precision and high
recall feature selection approaches.  We derive an efficient formula
for feature selection in this model using the filtering framework and
empirically compare to a wide variety of existing filtering-based
feature selection methods.  Results show that our high precision and
high recall feature selection approaches do reliably improve these
respective objectives.  Furthermore, our modeling framework provides
us with a way to interpret existing feature selection methods and
suggests that the use of greedy approaches (such as mutual
information) appear to maximize precision while diversity-based
approaches (such as MRMR) appear to maximize recall, hence providing
us with a framework for reinterpreting the loss functions implicitly
targeted by existing feature selection methods.