### 2014-01-31  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/abstract.tex
Abstract update

Feature selection is important in many practical classification
settings to address computational constraints or to prevent
overfitting when the data may be high-dimensional but contain
relatively few samples.  While a wide variety of feature selection
methods have been proposed in the literature, little research seems to
focus on feature selection specifically targeted for precision or
recall.  In this paper we propose a novel probabilistic model of
feature selection and argue that MAP inference w.r.t. to two extreme
objectives in this model respectively yield high precision and high
recall feature selection.  From each objective, we then derive an
efficient formula for feature selection in the filtering framework
(i.e., greedy forward selection) and we empirically compare the
resulting feature selection criteria to a wide variety of existing
methods.  Results show that our high precision and high recall feature
selection approaches improve their intended objectives and more
reliably than existing methods whose precision and recall performance
may vary widely across datasets and classifiers.

%Furthermore, our modeling framework provides us with a way to
%interpret existing feature selection methods and suggests that the
%use of greedy approaches (such as mutual information) appear to
%maximize precision while diversity-based approaches (such as MRMR)
%appear to maximize recall, hence providing us with a framework for
%reinterpreting the loss functions implicitly targeted by existing
%feature selection methods.

% could look at class balance
% what other feature properties?

% may want to mention why P/R important... depends on what is
% important (more costly for false positives or false negatives)
% and accuracy easy to optimize in imbalanced cases