### 2014-02-04 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/empirical.tex
Added Libo's related work

\section{Empirical}

In this section, we conduct experiments to study the feature selection
performances of several well known methods such as Symmetrical
Uncertainty Rank, Gain Ratio Rank, Information Gain Rank, Correlation
Based Rank, Conditional Entropy Rank \cite{guyon_jmlr03},
Correlation-based Feature Subset Selection \cite{Hall1998}, Reliff
\cite{Robnik-Sikonja2003} , mRMR \cite{peng2005} and our proposed
methods: High Precision Expectation, High Precision Log likelihood,
High Recall Expectation and High Recall Log likelihood. The first five
methods are variable rank methods and the last seven are subset rank
methods.

In these experimental study, we evaluated each feature selection
method in binary classification problems such as Breast Cancer, Pima
Diabetes, Heart Statlog, Spect and Vote. These data sets can be
downloaded from UCI machine learning repository
\cite{Bache+Lichman:2013}. In addition, we used three different
classifiers to solve each of the binary classifier problem. Logistic
Regression, SVM Linear and Naive Bayes were used in the experiments
and the first two are implemented in LibLinear library \cite{REF08a}
and the last one is implemented in Weka software \cite{weka}.

To evaluate the feature selection performances we reported
classification metrics performing 10-fold cross-validation from one
until the max number of features selected for each classifier solving
each of the binary classification problem. Specifically, we reported
Accuracy, Precision, Recall and F-Score for the true label class, that
could be identified by the context of the problem and usually be the
class with least samples in the data set.