### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/background.tex
Update

\section{Background}

in well-known contingency table

\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Label T} & \multicolumn{1}{l}{Label F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}

accuracy = tp + tn / tp + fp + fn + tn

precision = tp / tp + fp

recall = tp / tp + fn