### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/related_work.tex

\section{Related Work}

Existing feature selection algorithms fall into categories of filter
methods and wrapper methods \cite{guyon_jmlr03}. While
wrapper methods, such as SVM-RFE \cite{guyon2002gene}, select features
by evaluating their usefulness to a given classifier, filter methods
evaluate feature subsets according to certain properties of
themselves. Within filter methods, feature selection algorithms can be
further classified into ranking methods and subset methods. The former
evaluates each features independently, while the latter evaluates a
subset at a time \cite{brown2012conditional}. Listed below is a brief
introduction to algorithms to which we compared our proposed model.

Correlation based rank is a na\"{i}ve ranking algorithm that ranks
features according to their linear correlation with the output,
i.e.\ Pearson's $r.$ More sophisticated ranking methods include:
\begin{enumerate}
\item The conditional entropy of class $Y$ given feature $X,$
  i.e.\ $H\left(Y|X\right),$ quantifies the amount of information in
  $Y$ that is not provided by $X.$ Subtracting $H\left(Y|X\right)$
  from the entropy of class $Y,$ the information gain of class $Y$
  given feature $X$ is the mutual information between feature $X$ and
  class $Y,$ i.e.\ $I\left(X;Y\right).$
\item The gain ratio of feature $X$ is defined as the information gain
  of feature $X$ normalized against the entropy of itself,
  i.e.\ $\frac{I\left(X;Y\right)}{H\left(X\right)}.$
\item The symmetric uncertainty between feature $X$ and class $Y$
  measures the amount of redundancy between them. It is defined as
  $U\left(X,Y\right)=2\frac{I\left(X;Y\right)}{H\left(X\right)+H\left(Y\right)}.$
\item The Relief method \cite{kira1992feature} evaluates the relevance
  of features to the output class according to how well their values
  distinguish between nearest instances of the same and different
  classes.
\end{enumerate}

Correlation based subset \cite{Hall1998} is an extension
of Pearson's $r$ to subset methods. It measures the linear correlation
between each pair of selected features in addition to the correlation
between selected features and the output class. The other subset
algorithm to which we compared our work is MRMR \cite{peng2005}.
MRMR is an information-theory-based method that not only maximises
the relevance of selected features to the supervised output class,
but also minimises the redundancy among selected features. Different
search strategies can be applied using these two algorithms as
heuristics.

Although existing algorithms give various solutions, they are mostly
ad-hoc, until \cite{brown2012conditional}. Brown et al.\ derived an
information-theoretical model from first principles, and retro-fitted
existing information-theory-based feature selection methods to the
model. In this paper, we consider the other side of the same coin. We
proposed a probability-theory-based graphical model that has intuitive
qualitative meaning in precision and recall, and we built a
generalized feature selection scheme upon it.