### 2014-02-05  feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/background.tex
Current version

\section{Classification and Feature Selection}

In this section, we briefly define the task of binary classification
along with standard definitions of performance metrics we may wish 
to optimize.  We then follow this by a discussion of feature selection and
existing criteria proposed in the literature.

In the binary classification task, we assume we are given data $D = \{
(\vec{x}^d, y^d) \}$ consisting of pairs of real-valued raw 
feature vectors $\vec{x}^d \in \mathbb{R}^n$
of length $n$ (e.g., counts of $n$ different words in a document) and
\emph{actual} binary class label $y^d \in \{ 0 (\mathrm{false}), 1 (\mathrm{true}) \}$ (where
we often write F for false and T for true).
A binary classifier is a function $C: \vec{x}^d \to y^d$ such that given
a new unlabeled raw feature vector, $C(\vec{x}^d)$ produces a \emph{predicted}
classification.

Given a trained classifier $C$ and a dataset $D$, we can build the 
well-known contingency table
\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Actual T} & \multicolumn{1}{l}{Actual F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}
where each of the four entries

\begin{align*}
\textrm{Accuracy}  & = \frac{TP + TN}{TP + FP + FN + TN}
\textrm{Precision} & = \frac{TP}{TP + FP}
\textrm{Recall}    & = \frac{TP}{TP + FN}
\end{align*}