### 2014-02-05 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/conclusions.tex
Update

\section{Conclusion}

In this paper we argued that no feature selection methods have
previously targeted recall for the task of binary classification and
we proposed a model and objectives to address this task.  Empirically,
our results show that our high recall feature selection approach shows
strong performance on recall with significant improvements over other
methods in cases where there is a high feature to data ratio.
Future extensions of this work may examine tighter integration of the
feature selection method with the algorithm to make an \emph{embedded}
approach 
using the classifier implicit in our model.
%based on our model since at it's core is a voting-based
%classifier.