### 2014-02-05 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/background.tex
Update

\section{Classification and Feature Selection}

In this section, we briefly define the task of binary classification
along with standard definitions of performance metrics we may wish 
to optimize.  We then follow this by a discussion of feature selection.
% and existing criteria proposed in the literature.

In the binary classification task, we assume we are given data $D = \{
(\vec{x}^d, y^d) \}$ consisting of pairs of real-valued raw 
input vectors $\vec{x}^d \in \mathbb{R}^n$
of length $n$ (e.g., the results of $n$ different medical tests) and
\emph{actual} binary class label $y^d \in \{ 0 (\mathrm{ false}), 1 (\mathrm{ true}) \}$ (where
we often write F for false and T for true).
A binary classifier is a function $C: \vec{x}^d \to y^d$ such that given
a new unlabeled raw feature vector, $C(\vec{x}^d)$ produces a \emph{predicted}
classification.

Given a trained classifier $C$ and a dataset $D$, we can build the 
well-known contingency table
\begin{center}
\begin{tabular}{l|l|l|} 
\multicolumn{1}{l}{} &  \multicolumn{1}{l}{Actual T} & \multicolumn{1}{l}{Actual F} \\ \cline{2-3}
Predicted T & TP & FP \\ \cline{2-3}
Predicted F & FN & TN \\ \cline{2-3}
%Predicted T & TP & FP & \#$T_P$ \\ \cline{2-3}
%Predicted F & FN & TN & \#$F_P$ \\ \cline{2-3}
%          &   & \#$T_A$ & \#$F_A$ & \\ \cline{2-3}
\end{tabular}
\end{center}
where the four entries represent the counts of true positives (TP), false positives (FP), false negatives (FN)
and true negatives (TN) and sum to the 
total amount of data (i.e., $\mathrm{TP} + \mathrm{FP} + \mathrm{FN} + \mathrm{TN} = |D|$).
Each table entry represents the count of data for which the respective row matched the 
predicted classification $C(\vec{x}^d)$ and the respective column matched the actual label $y^d$.
Given these definitions, we can easily define 
\begin{align*}
\textrm{Accuracy}  & = \frac{TP + TN}{TP + FP + FN + TN} \\
%\textrm{Precision} & = \frac{TP}{TP + FP}
\textrm{Recall}    & = \frac{TP}{TP + FN}
\end{align*}
where accuracy represents the overall fraction of correct classification, but recall
represents the overall fraction of true labeled data that has been classified as
true (i.e., recalled).  For somewhat rare events such as medical diagnosis of cancer,
we certainly care about accuracy, but we may also want to place additional emphasis
on recall performance so as to avoid the occurrence of false negatives (cases of cancer
that were missed by the classifier).  Of course false positives are also a problem,
but additional tests would rule these out and hence not as critical of a classification
failure as missing a potential cancer diagnosis.

Previously we did not specify exactly how $C(\vec{x}^d)$ learns from the raw data
$\vec{x}^d$.  In general, practitioners often process the raw input vectors
$\vec{x}^d$ into a subset of features that we'll denote $F_k = \{ f_1,\ldots,f_k \}$.  Whereas
the raw input may consist of the results of individual tests, a feature $f_i \in \mathbb{R}$
may represent some nonlinear function of one or more tests deemed to be relevant to
the classification task.  Hence we might more appropriately write a classifier
as $C(\vec{x}^d,F_k)$ to represent the raw input data and the features
of the data that the classifier may use.
In the case of high-dimensional data or few data samples,
we may wish to limit the set of features generated to improve classifier performance
and this is the task of feature selection --- select $F_k$ to optimize
performance.  There are already a variety of feature selection methods that we outline
next.