### 2014-02-05 trunk/ feature-selection-high-recall/FeatureSelectionProject/doc/fs_hr/conclusions.tex
Final version

\section{Conclusion}

In this paper we argued that no feature selection methods have
previously targeted recall for the task of binary classification and
we proposed a model and objectives to address this task.  Empirically,
our results show that one of our novel feature selection methods (CVL)
shows strong performance on recall with significant improvements over
other methods in cases where there is a high feature to data ratio.
Future work may examine tighter integration of our feature selection
method with its implicit classification algorithm in an efficient
embedded approach to high-recall feature selection.
%using the classifier implicit in our model.
%based on our model since at it's core is a voting-based
%classifier.